---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---


## Loading and preprocessing the data

* Download zip file if not alredy present
* Unzip if if not present
* Load Data


```{r}
archiveFile <- "activity_data.zip"
if(!file.exists(archiveFile)) {
    archiveURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    download.file(url=archiveURL,destfile=archiveFile,method="curl")
}

if(!(file.exists("activity.csv"))) { unzip(archiveFile) }

activity = read.csv("activity.csv")

```

Summary of activity data

```{r}

summary(activity)

```


## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r}

totalSteps <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)

summary(totalSteps)

```


 Make a histogram of the total number of steps taken each day

```{r}
hist(totalSteps$steps)
``` 

Calculate and report the mean and median of the total number of steps taken per day

```{r results="hide"}
options(scipen = 1, digits = 2) #Format output 
meanSteps = mean(totalSteps$steps)

medianSteps = median(totalSteps$steps)

```

Mean steps is `r meanSteps` and median steps is `r medianSteps`

## What is the average daily activity pattern?

Make a time series plot

```{r}
stepsInterval <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(steps ~ interval, data = stepsInterval, type = "l")

```


```{r}
maxStepsInterval = stepsInterval[which.max(stepsInterval$steps), ]$interval
```

Max steps occured in the `r maxStepsInterval` interval

## Imputing missing values

Total number of missing values 

```{r}
sum(is.na(activity$steps))

```


```{r}
interval2steps <- function(interval) {
    stepsInterval[stepsInterval$interval == interval, ]$steps
}

activityFilled <- activity  
count = 0  
for (i in 1:nrow(activityFilled)) {
    if (is.na(activityFilled[i, ]$steps)) {
        activityFilled[i, ]$steps <- interval2steps(activityFilled[i, ]$interval)
        count = count + 1
    }
}


```

A total of `r count` valus were filled.

Make a histogram of the total number of steps with filled data

```{r}
totalSteps2 <- aggregate(steps ~ date, data = activityFilled, sum)
hist(totalSteps2$steps)

```

Calculate and report the mean and median

```{r}
meanStepsFilled = mean(totalSteps2$steps)

medianStepsFilled = median(totalSteps2$steps)

```


Mean steps is `r meanStepsFilled` and median steps is `r medianStepsFilled`

Do these values differ from the estimates from the first part of the assignment? 

Mean values are the same since mean was used to fill in data.
 
What is the impact of imputing missing data on the estimates of the total daily number of steps?

The median value shows a slight change up.



## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend”

```{r}
activityFilled$day = ifelse(as.POSIXlt(as.Date(activityFilled$date))$wday%%6 == 
    0, "weekend", "weekday")
# For Sunday and Saturday : weekend, Other days : weekday
activityFilled$day = factor(activityFilled$day, levels = c("weekday", "weekend"))

```

Make a panel plot containing a time series plot 

```{r}

stepsInterval2 = aggregate(steps ~ interval + day, activityFilled, mean)
library(lattice)
xyplot(steps ~ interval | factor(day), data = stepsInterval2, aspect = 1/2, 
    type = "l")
```


Weekend activity appears to be overall higher, while weekday activity has a spike in the A.M. and generally lower as the day moves on.